---
title: "Gov 50 Final Project"
author: Lauren Alexes Kirkpatrick
description: "Money as a Means to Happiness: A Start or an End?"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction
Philosophy. How should we be living? 
 

## Data Section: Source, key dependent & independent variables 


```{r}
library(tidyverse)
library(dplyr)
Data1 <- read_csv("WHR2023.csv")
colnames(Data1)

WHDF2023 <- Data1 |>
  select(-c(`Ladder score in Dystopia`:`Dystopia + residual`)) |>
  select(-c(`Standard error of ladder score`: `lowerwhisker`)) |>
  rename(country = `Country name`, 
         happiness_score = `Ladder score`, 
         GDP_per_capita =`Logged GDP per capita`,
         social_support = `Social support`,
         life_exp =  `Healthy life expectancy` , 
         freedom =  `Freedom to make life choices` ,
         generosity = `Generosity`, 
         corruption = `Perceptions of corruption`) |>
  arrange(desc(happiness_score))
WHDF2023$Year <- "2023"
WHDF2023


Bottom_WHDF2023 <- WHDF2023 |>
    slice_min(happiness_score, n = 5)
Bottom_WHDF2023

Top_WHDF2023 <- WHDF2023 |>
    slice_max(happiness_score, n =5) 
Top_WHDF2023




WHDF2023_10 <- rbind(Top_WHDF2023, 
                     Bottom_WHDF2023)
WHDF2023_10

```


```{r}
Data2 <- read_csv("WHR2019.csv")
WHDF2019 <- Data2 |>
  select(-c(`Overall rank`)) |>
  rename(country = `Country or region`, 
         happiness_score = Score, 
         GDP_per_capita =`GDP per capita`,
         social_support = `Social support`,
         life_exp =  `Healthy life expectancy` , 
         freedom =  `Freedom to make life choices` ,
         generosity = `Generosity`, 
         corruption = `Perceptions of corruption`) |>
  arrange(desc(happiness_score)) 


WHDF2019$Year <- "2019" 
WHDF2019

WHDF2019_10 <- 
  WHDF2019[c(1,2,4,13,5,154,91,129,146,127),]

WHDF2019_10
```


## A Summary of the Dependent Variable (The Main Outcome of Interest)

```{r}
WHDF2019_VS_WHDF2023 <- rbind(WHDF2019_10, 
                     WHDF2023_10)
WHDF2019_VS_WHDF2023
```
```{r}
plot1 <- ggplot(data = WHDF2023 , 
       mapping = aes(x = GDP_per_capita, 
                     y = happiness_score)) +
         geom_point() +
         geom_smooth(lm = loess) 
      

plot1
```
```{r}
plot2 <- ggplot(data = WHDF2023 , 
       mapping = aes(x = GDP_per_capita, 
                     y = happiness_score)) +
         geom_point(color = "orange") +
         geom_smooth(color = "gold") +
   scale_x_log10()
      

plot2
```
```{r}
plot2 <- ggplot(data = WHDF2023 , 
       mapping = aes(x = social_support, 
                     y = happiness_score)) +
         geom_point(color = "orange") +
         geom_smooth(color = "gold") +
   scale_x_log10()
      

plot2
```
```{r}
plot3 <- ggplot(data = WHDF2023 , 
       mapping = aes(x = generosity, 
                     y = happiness_score)) +
         geom_point(color = "orange") +
         geom_smooth(color = "gold") +
   scale_x_log10()
      

plot3
```
## Results: Main Analysis & Regression Output 

```{r}
mean_happ_7 <- WHDF2023  |>
  filter(round(happiness_score) == 7) |>
  summarize(mean(GDP_per_capita)) |>
  pull()
mean_happ_7
```
```{r}
mean_happ_7_2023 <- WHDF2023_10 |>
  filter(round(happiness_score) == 7) |>
  summarize(mean(GDP_per_capita)) |>
  pull()
mean_happ_7_2023
```
```{r}
mean_happ_7_gen <- WHDF2019_10 |>
  filter(round(happiness_score) == 7) |>
  summarize(mean(generosity)) |>
  pull()
mean_happ_7_gen
```

```{r}
mean_happ_7_gen_2023 <- WHDF2023_10 |>
  filter(round(happiness_score) == 7) |>
  summarize(mean(generosity)) |>
  pull()
mean_happ_7_gen_2023
```


```{r}
library(broom)
ggplot(data = WHDF2023 , aes(x = GDP_per_capita, y = happiness_score)) + geom_point(size = 1.5) +
  
 geom_point(aes(x = mean_happ_7, y = 7), color = "lightgoldenrod", size = 4)
             

```
```{r}
ggplot(data = WHDF2023 , aes(x = GDP_per_capita, y = happiness_score)) + geom_point(size = 1.5) + 
  stat_summary_bin(fun = "mean", color = "lightgoldenrod", size = 4, geom = "point", binwidth = 1) + 
  geom_smooth(method = "lm", se = FALSE, color = "gold", size = 1.5)
  
```
```{r}
ggplot(data = WHDF2023_10, aes(x = happiness_score, y = GDP_per_capita, color = country)) + geom_point(size = 1.5) + 
  stat_summary_bin(fun = "mean", color = "lightgoldenrod", size = 4, geom = "point", binwidth = 1) + 
  geom_smooth(method = "lm", se = FALSE, color = "gold", size = 1.5)
```

```{r}
ggplot(data = WHDF2019_10, aes(x = generosity, y = happiness_score, color = country)) + geom_point(size = 1.5) + 
  stat_summary_bin(fun = "mean", color = "lightgoldenrod", size = 4, geom = "point", binwidth = 1) + 
  geom_smooth(method = "lm", se = FALSE, color = "gold", size = 1.5)
```
```{r}
ggplot(data = WHDF2023_10, aes(x = generosity, y = happiness_score, color = country)) + geom_point(size = 1.5) + 
  stat_summary_bin(fun = "mean", color = "lightgoldenrod", size = 4, geom = "point", binwidth = 1) + 
  geom_smooth(method = "lm", se = FALSE, color = "gold", size = 1.5)
```

## Hypothesis Testing...
Use findings to reject the hypothesis that a change in GDP does not change the score? We reject the null hypothesis if p is greater than alpha value, which will be .05. 


```{r}
fit <- lm(happiness_score ~ GDP_per_capita + social_support + generosity, data = WHDF2023 )

broom::tidy(fit) |> knitr::kable()  
```
```{r}
generosity_happiness_score_lm <- lm(happiness_score ~ generosity, data = WHDF2019_VS_WHDF2023)

summary(generosity_happiness_score_lm)
```
```{r}
plot(generosity_happiness_score_lm)
```





