---
title: "Gov 50 Final Project"
author: Lauren Alexes Kirkpatrick
description: "Money as a Means to Happiness: A Start or an End?"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction: Research Question and Main Hypothesis 
The following project aims to explore the universal strive towards happiness in a world that indeed makes this endeavor seem nearly impossible. What particular factors of one's life and state significantly impact one's happiness? In particular, does one's economic standing play a greater role in one's happiness compared to more social factors, one's relationships? This projects hopes to debunk the notion that "money makes you happy." Thus, I hypothesize that indeed one's relationships more significantly influence one's happiness compared to that of one's economic standing, money.

Certainly, one cannot control the happiness of another. However, governments still have an obligation to consider the happiness of its citizens. After all, it is governments who are in a position of power to pursue the necessary means to enact such state-wide change. Therefore, through identifying what exactly drives one's happiness, governments can better formulate initiatives to ensure the achievement of happiness for its citizens, more specifically the accessibility of this universal right. This is to say that the following research question is important as it guides governments on how to address humanity's greater strive towards happiness and universal right to achieving happiness. 

## Data Section: Source, Measurement of Independent and Dependent Variables, Research Design

To answer this question of what factor influences one's happiness the most, this research project will be using data from the World Happiness Report for the year of 2023. This report stems from the Sustainable Development Solutions Network, which is driven by the Gallup World Poll data. More importantly, the following organization similarly aims to explore happiness with the goal of calling upon governments to consider their own roles in the happiness of its citizens. The data of the report was collected through evaluations made by respondents regarding their own perception of their happiness on the "Cantril ladder." In the context of this source and thus this report, happiness is defined as one's "life evaluation." The following scale asked respondents to think of a ladder with 10 representing one's "best possible life" and 0 representing the worst. They ultimately rated their own lives on that scale. This information was then observed on six variables (social support, healthy life expectancy, freedom, generosity, corruption, and GDP per capita). Therefore, the happiness rankings that are earned by countries are not based on these factors, but rather the individual assessments of respondents on their lives (i.e. their answer to the Cantril ladder evaluation). (https://worldhappiness.report/about/)

The Gallup World Poll data, the main source of the World Happiness Report, is a survey based data collecting tool. According to the Gallup agency, this poll is conducted through a combination of telephone surveying along with in-person interviewing. For countries that have telephone coverage for at least 80% of the population, telephone surveys are leveraged. Additionally, where telephone interviewing is employed, a random-digit-dial is used. In countries that do not have the above telephone coverage, households are randomly selected for in-person interviewing. Therefore, the RESEARCH DESIGN of this data is a randomized experiment. (https://www.gallup.com/178667/gallup-world-poll-work.aspx)

The above research question can be simplified to exploring the influence of GDP per capita and one's relationships (social support) on one's happiness. Therefore, the main outcome of interest is happiness score, our dependent variable. The independent variables in this case, are a country's GDP per capita and its score of social support. 


```{r}
library(tidyverse)
library(dplyr)
library(kableExtra)
Data1 <- read_csv("WHR2023.csv")
colnames(Data1)

WHDF2023 <- Data1 |>
  select(-c(`Ladder score in Dystopia`:`Dystopia + residual`)) |>
  select(-c(`Standard error of ladder score`: `lowerwhisker`)) |>
  select(-c(`Healthy life expectancy`:`Perceptions of corruption`)) |>
  rename(country = `Country name`, 
         happiness_score = `Ladder score`, 
         GDP_per_capita =`Logged GDP per capita`,
         social_support = `Social support`) |>
  arrange(desc(happiness_score)) 
WHDF2023



WHDF2023 |>
  kbl(caption = "World Happiness Data 2023", col.names = c("Country", "Happiness Score", "GDP per capita", "Social Support")) |>
  kable_classic(full_width = F, html_font = "Cambria")

```


```{r}
plot1 <- ggplot(data = WHDF2023, 
       mapping = aes(x = GDP_per_capita, 
                     y = happiness_score)) +
         geom_point() +
         geom_smooth(lm = loess) 
plot1
```
```{r}
plot2 <- ggplot(data = WHDF2023, 
       mapping = aes(x = social_support, 
                     y = happiness_score)) +
         geom_point() +
         geom_smooth(lm = loess) 
plot2
```


## Results: Main Analysis & Regression Output 

```{r}
mean_happ_7 <- WHDF2023 |>
  filter(round(happiness_score) == 7) |>
  summarize(mean(GDP_per_capita)) |>
  pull()
mean_happ_7
```
```{r}
mean_happ_7_ss <- WHDF2023 |>
  filter(round(happiness_score) == 7) |>
  summarize(mean(social_support)) |>
  pull()
mean_happ_7_ss
```

```{r}
library(broom)
ggplot(data = WHDF2023 , aes(x = GDP_per_capita, y = happiness_score)) + geom_point(size = 1.5) +
  
 geom_point(aes(x = mean_happ_7, y = 7), color = "lightgoldenrod", size = 4)
             
```
```{r}
ggplot(data = WHDF2023 , aes(x = social_support, y = happiness_score)) + geom_point(size = 1.5) +
  
 geom_point(aes(x = mean_happ_7_ss, y = 7), color = "lightgoldenrod", size = 4)
             
```

## Hypothesis Testing...
Use findings to reject the hypothesis that a change in GDP does not change the score? We reject the null hypothesis if p is greater than alpha value, which will be .05. 

```{r}
fit <- lm(happiness_score ~ GDP_per_capita + social_support, data = WHDF2023 )

broom::tidy(fit) |>  
  kbl(caption = "Hypothesis Testing", col.names = c("Independent Variable", "Estimate", "Standard Error", "Statistic", "P Value")) |>
  kable_classic(full_width = F, html_font = "Cambria")
```








